A Microservices Based Approach for City Traffic Simulation

### 2019-12-08 papers/traffic simulation/paper.tex
Some corrections.

Simulating traffic in a city is a difficult task. Current paper is proposing a novel way using state of the art in distributed systems: microservices orchestration. The design is centered around two types of actors: a city simulation actor (a single instance of it) which keeps track of occupied streets and and their gradual occupation. The other actor type (multiple instances) is the car which travels across the city.

@@

A solution which simulates car, pedestrian, etc. traffic in a given city may reap many benefits. It can help understand patterns of traffic and its flow. It can help unerstand the particularities and pecularities of a city's streets arrangements. It can help identify bottlenecks. It can help find solutions to rush problems and explore them. But for those areas to be tackled, appropiate methods of simulating traffic must be found. We are proposing a novel way of simulating traffic based on the microservices orchestration concept and using a discrete microscopic logic.

We define a \textbf{(city) actor} as being a independent entity which chose to move between two geographical points within a city. It can be a car, a pedestrian or a bike. We also define the \textbf{city simulator} as being a single entity (subject to distributed and load balacing services) which keeps data about the city (e.g. streets with city actors on them).

This paper is structured in few sections. Section \ref{sec:existingsolutions} contains a summary of few other papers and ideas around the same subject of traffic simulatino. Section \ref{sec:implementation} discuss the core approach of the current paper in architecting a software solution based on microservices for simulating traffic in a city. Section \ref{sec:orchestration} discuss the architectural aspect of orchestrating microservices using popular frameworks and solutions. Section \ref{sec:results} presents the results obtained after the implementation of some of the core concepts from the previous sections.

@@

\cite{6973934} is describing a multiagent system for interactions between pedestrians and vehicles, where each pedestrian and vehicle is an intelligent agent with own personality and decisions. This idea is an interesting one especially because it is proposed as an enhacement to the solution presented by the current paper. In the same area of microscopic simulation we can see another proposal from \cite{7004985}. Of a great interest is their ability to feed back in the simulation the data gathered from loop detectors across the portion of highway they are particularly interested in.

@@

In \cite{8569576} the broad idea is to us a discrete event architecture where there are logical processes for executing a number of agents. For a computation intensive those processes can be clustered into agent clusters with dependencies between them. The particularity of this proposal is that it crosses the boundaries of a single computing machine and pave a way to distribute the load and information across machines, with certain limitations and challenges.

@@

Thus, a container is a small set of processes which run in isolation. They allow packaging a linux distro, a set of libraries and sdk and on top of those custom code. This forms an image, which is essentially a tar gzipped file. Once the build process of an image is finished, it can be spinned up in one or more running containers. The custom code written and embedded in the image is running in parallel in each container.

@@

The most common and easy solution to share data across all the city actors is to have a centralized store to keep it. City simulator acts as a centralized store for all other city actors. In the current implementation the city simulator keeps a set of data which can be described as a list of pairs, each pair having a line and a real number, called density.

@@

Figure \ref{fig:citysimrelations} depicts the city simulator in relation with the other entities. A notable exception is the web page. Its purpose is not to participate in the same information exchange the other entities have but to offer a visual interpretation of the way the city and the other actors are interacting with one another.

@@

The city actor represents a moving actor within a city. It can be a car which moves across the city, it can be a bike or it can be a pedestrian. Its naming suggests that it can be any entity or living being which moves within a city and interacts with the other entities or affects the other entities in some manner. A pedestrian would directly interact with other pedestrians but not with cars. A pedestrian would indirectly affect other cars by willing to walk over a street crossing.

The implementation proposed here is aiming to model a single type of actor: a car which moves between two points across a city. Because the purpose of this paper is not to tackle maps representation and routing through a city (in itself, this area is way bigger than a mere technical paper) the city actor is using two notable services: Open Street Map \citep{openstreetmap} and GraphHopper \citep{graphhopper}. Open Street Map is an open source licensed map of the entire world. Graphhoper is an open source service which offers directions APIs and route planning. It can use Open Street Map as an underlying map provider. They provider free tiers and paid subscription for accessing an API and get routes. However, a city actor is not using any public api but a special crafted container which contains GraphHopper and an open street map embedded into it. Thus, this container is running in parallel with the other city actors and provides them with the routing API they need.

@@

An actor interacts, for the time being, only with the city simulator by using a Go struct and a Go enum. They can be seen in listings \ref{lst:goactorstruct} and \ref{lst:goactorreports}. The message sent from one side to another has a general structure called \textit{Envelope}. It is meant to offer a top level message which can be serialized (or encoded, the way to do this in Golgang, if not gRPC, is the gob pacakge it offers out of the box) and passed around, enabling any party involved to understand what this message is about and how to decode it.

@@

\textit{AskForLine} is a message sent from an actor to the city simulator in which the actor asks about the density of any line. This way any actor can take conscious decisions on which route to go on when based on what lies ahead in terms of street densities. When a first route is known, between two desired points, the actor can ask about each line it has on that route. The city simulator will respond with the known density of it. If the actor desires, it can try to find another route by asking GraphHopper service to compute a new one but with an additional rule: avoid a certain point (street, junction, etc.). \textit{RespondWithLine} is the type of message which the city simulator sends back to an actor after it received an \textit{AskForLine} message.

@@

As with every software project started from scratch there are a number of choices to make when choosing software stacks, programming languages, networking models, etc. This section is aiming to explore the rationale behind some of those decisions and how they influenced the building of  such an application.

@@

Go routines are a lightweight thread of executing. Being lightweight it means they can be easily started, with no overhead and especially without the ceremony of dealing with threads which other popular languages (e.g. Java) have. Listing \ref{lst:goroutine} shows such a routine. It is a part of logic where the city actor responds to either two events: the timer has expired and a decision has to be made or an information about the line is currently traveling on has been received from the city simulator. Another Go trait is the channels, an indexed communication pipe where there are asynchronous writers and readers, used to decouple two routines. Listing \ref{lst:goroutine} also shows couple of channels. Variables \textit{ticker} and \textit{lineChan} are two channels. They are declared using a type and after that they are used to write and read objects (or structs) of the respective type. Combined with the \textit{select} keyword, it becomes easy to write code which takes advantage of multithreading paradigms.

@@

The term orchestration means the handling of containers and microservices in order to bring coherence into their interaction and an unifying experience to the end user of a service or of a product. They migth run on multiple nodes (computers or virtual machines) and at the same time they need to communicate with one another. They need to be updated in place and without service disruption. Whenever one of them crash they have to restart as quickly as possible. They have to be able to discover themselves without dealing with intricancies of IP addresses, proxies, etc. Those are just few concerns around orchestration and the current paper is not aiming to provide a comprehensive view of what it means and what can be achieved with it but rather to set a basis of understanding enough context for the subject of simulating a city with its traffic.

@@

The city simulator is, currently, a single service which means it runs as a single instance. The city actor container runs in multiple instances and it has the need to do so as part of the entire simulation workload. The other two instances which run in the simulation are the routing service, a Graphhopper instance, and the front end web erver which displays a web page for offering visual clues around the simulation. While they run as such there should be an easy way, without friction and additional compute logic to "discover" a certain service. A city actor will need to create connect itself to the routing service and to the city simulator. At the same time the front end instance need to connect itself to the city simulator to source its data.

@@

The \textit{docker-compose} tool gives the ability to manipulate more containers and services, with a simple file written in yaml format. Listing \ref{lst:dockercompose} shows the docker-compose file in its brevity and briefness. Let's dissect it. There are four services: \textit{graphhopper}, \textit{citysim}, \textit{cityactor} and \textit{cityfront}. Each of them need an image to run and this image can be specified in two ways: either as an image already compiled and hosted on an container registry or as a local folder which contains an file to build one (defined by the presence of a file called \textit{dockerfile}). Because Graphhoper, once compiled with desired maps and settings do not need any more development work, it is taken as an image from a docker registry. The other three services are the places where the most development efforts take place therefore they need to be compiled or recompiled each time the entire traffic simulation application starts.

The containers which are running may need to have different properties. First and most important is the container port which needs to be published. The port from inside the container, where a certain process expect a TCP communication is forwarded to the local host port and thus is accessible from outside the Docker internal network. Another detail of the container is its dependency. For example \textit{cityactor} cannot run without textit{Graphhopper} because it doesn't have any place to obtain a route. Therefore it depends on \textit{Graphhopper}. And on \textit{citysim}, of course. The last bit of detail which can be seen here is the policy of restarting a container. By default the policy is set to "No" which means that the container will not be restarted if there is any failure within it and it stops. However, for simulation purposes, as we need a constant stream of city actors to swarm through the city, the policy is set to "Always" which will restart the container after if closes itself, i.e. it finishes its travel.

@@

Enter the DNS (Domain Name System), the backbone of the entire internet. In its simplest description, avoiding many inherent and nitpicking details, it is a dictionary which keeps track of every registered and easy to memorize name, e.g. en.wikipedia.com and the IP addresses where it is located. Any client which would like to access such an name will query first the DNS resolvers and then it will proceed to send a message to the obtain IP address. The most notable interaction of a user with the DNS is the address bar of a browser where the user inserts the name of the site they want to access and while writing the address suggestions are displayed by the browser with the help of recommandation engines \citep{risley2001domain}.

When many resources are needed, to serve a great number of users or to support a great number of city actors, a certain service might be so busy with serving data that the compute and memory resources it needs are greater than the hardware is able to give. Therefore it might be located at few addresses at the same time and any client which wants to access it should access the address where there are available compute and memory resources to serve its queries. Ideal would be to abstract or to decouple this information from the client and make it transparent for it. To do this the client needs to know only the service name and a certain "networking" entity should route its request to the appropiate available service. Such a entity could be a DNS authoritative server which, based on the load, availability and latency of the services will route the request to the appropiate node \citep{swildens2006scalable}.

Docker and Kubernetes are doing a similar job. They offer a DNS service and according to the load of each node where a service run, a request is rooted to a node which will be able to respond to it. This logic is completely decoupled from the clients. Listings \ref{lst:gocityclientdial} and \ref{lst:cityfrontclientdial} shows how both the city front and a city actor calls the city simulator. All what they do is to use the address which was specified in the docker-compose file (listing \ref{lst:dockercompose}). Docker or Kubernetes will do the actual job to route the request to the appropiate node of the service. One thing to note is the difference between the two. A city client will communicate via TCP/IP using a Go language Glob format (add referenc to the design choices paragraph) while the city front use a WebSocket communication technology (add reference to the design choices paragraph) but both of them are transparently routed by Docker.

@@

Figure \ref{fig:ScreenshotWebInterface} shows a screenshot of the web interface discussed in paragraph \ref{subsection:citysimulator}, showing a simulation with less than 20 actors. The web interface is a simple, httml/js only web page (vanilla js) which connects to the city simulator as can be seen in listing \ref{lst:cityfrontclientdial}. The city simulator will send a notification the web page server each time the density changes for a certain line. It can be that a new or first actor entered a certain line (a list of coordinates which represent a street) or it can be that one or the last actor left a line. The web page will draw any line with a density greater than 0 on top of the map.

@@

A laptop used for simulating (MacBook Pro, 16 GB RAM, i5 3.1 GHz) is able to run easily 50 city actors. Above a certain threshold the bottleneck, surprisingly, is not the memory or the cpu but the way docker handles network interfaces. For a larger number of ciy actors a better tool must be used (see paragraph \ref{subsec:kubernetes}). Figure \ref{fig:ScreenshotWebInterface2} shows a simulation with 50 actors running at the same time through the city.

@@

As with all technicl projects, there are a number of enhacements which can be further developed around the concept of simulating a city traffic through actors running in microservices, as presented in this paper. This section will briefly propose few of those details in the following paragraphs.

@@

A second idea for enhancing the simulation is to make the actors to have a more granular and diverse logic of traversing the city. The solution proposed into this paper is a basic one: an actor has a route between two random points on the map and then it goes on to travel across that route. It can also be made to look for street densities in advance and act accordingly. If a street is too crowded the actor might choose to avoid it and compute a new route around it.

A third idea is to add more interactions between actors. It has been discussed that a first interaction is and indirect one: the number of actors on a certain street, number called density. But there are many more points of interactions. Let's take junctions. The city simulator can model any junction and then it can permit actors to pass or not through it. Then there are pedestrian crossings where different types of actors can meet each other. Or subways access, elevators for them, bus stations, etc.

A fourth idea is to have contained actors. A bus is an actor. A person is an actor. If a person is waiting on a bus station and then it takes a bus, there are two actors, but they are tied and are not independent anymore until the person choose to get out of the bus. This means the densities across sidewalkings and bus stations need to be computed accordingly.

A fifth idea of enhancement is not tied to the way simulation is working but to how it is presented. The web page can display different colors based on the density of a line, offering a visual clue on it. Also it can easily show more information when the cursor is over a point of interest (a street, a junction, etc.).

>>>>

Simulating traffic in a city is a difficult task. Current paper is proposing a novel way using the state of the art in today's distributed systems: microservices orchestration. The design is centered around two types of actors: a city simulation actor (a single instance of it) which keeps track of occupied streets and and their gradual occupation. The other actor type (multiple instances) is the car which travels across the city.

@@

A solution which simulates car, pedestrian, etc. traffic in a given city may reap many benefits. It can help understand patterns of traffic and its flow. It can help understand the particularities and pecularities of a city's streets arrangements, together with their junctions. It can help identify bottlenecks. It can help find solutions to rush problems and explore them. But for those areas to be tackled, appropiate methods of simulating traffic must be found and explored. We are proposing a novel way of simulating traffic based on the microservices orchestration concept and using a discrete microscopic logic for the behavior of each actor.

We define a \textbf{(city) actor} as being an independent entity which chose to move between two geographical points within a city. It can be a car, a pedestrian or a bike. We also define the \textbf{city simulator} as being a single entity (subject to distributed and load balacing services) which keeps data about the city (e.g. streets with city actors on them).

This paper is structured in few sections. Section \ref{sec:existingsolutions} contains a summary of few other papers and ideas around the same subject of traffic simulation. Section \ref{sec:implementation} discuss the core approach of the current paper in architecting a software solution based on microservices for simulating traffic in a city. Section \ref{sec:orchestration} discuss the architectural aspect of orchestrating microservices using popular frameworks and solutions. Section \ref{sec:results} presents the results obtained after the implementation of some of the core concepts from the previous sections.

@@

\cite{6973934} is describing a multiagent system for interactions between pedestrians and vehicles, where each pedestrian and vehicle is an intelligent agent with its own personality and decisions. This idea is an interesting one especially because it is proposed as an enhancement to the solution presented by the current paper. In the same area of microscopic simulation we can see another proposal from \cite{7004985}. Of a great interest is their ability to feed back in the simulation the data gathered from loop detectors across the portion of highway they are particularly interested in.

@@

In \cite{8569576} the broad idea is to use a discrete event architecture, in which there are logical processes for executing and simulating a number of agents. For a computation intensive aim those processes can be clustered into agent clusters with dependencies between them. The particularity of this proposal is that it crosses the boundaries of a single computing machine and pave a way to distribute the load and information across machines, with certain limitations and challenges.

@@

Thus, a container is a small set of processes which run in isolation. They allow packaging a linux distro, a set of libraries and a software development kit and on top of those the custom code of an application. Taken together, those form an image, which is essentially a tar gzipped file. Once the build process of an image is finished, it can be spinned up in one or more running containers. The custom code written and embedded in the image is running in parallel in each container.

@@

The most common and easy solution to share data across all the city actors is to have a centralized store to keep it. City simulator acts as a centralized store for all other city actors. In the current implementation the city simulator keeps a set of data which can be described as a list of pairs, each pair having a line and a real number, called density. The line is a series of coordinates and in the proposed implementation their meaning is a street which a city actor is reportedly traveling on it.

@@

Figure \ref{fig:citysimrelations} depicts the city simulator in relation with the other entities. A notable exception is the web page. Its purpose is not to participate in the same information exchange the other entities have but to offer a visual interpretation about the way the city and the other actors are interacting with one another.

@@

The city actor represents a moving actor within a city. It can be a car which moves across the city, it can be a bike or it can be a pedestrian. Its naming suggests that it can be any entity or living being which moves within a city and interacts with the other entities or affects the other entities in some manner. A pedestrian would directly interact with other pedestrians but not with cars unless it crosses the street on red or on unmarked places. A pedestrian would indirectly affect other cars by willing to walk over a street crossing.

The implementation proposed here is aiming to model a single type of actor: a car which moves between two points across a city. Because the purpose of this paper is not to tackle maps representation and routing through a city (in itself, this area is way bigger than a mere technical paper) the city actor is using two notable services: Open Street Map \citep{openstreetmap} and GraphHopper \citep{graphhopper}. Open Street Map is an open source licensed map of the entire world. Graphhoper is an open source service which offers directions APIs and route planning. It can use Open Street Map as an underlying map provider. They provide free tiers and paid subscription for accessing an API and compute various routes. However, a city actor is not using any public api but a special crafted container which contains GraphHopper and an open street map embedded into it. Thus, this container is running in parallel with the other city actors and provides them with the routing API they need.

@@

An actor interacts, for the time being, only with the city simulator by using a Go struct and a Go enum. They can be seen in listings \ref{lst:goactorstruct} and \ref{lst:goactorreports}. The message sent from one side to another has a general structure called \textit{Envelope}. It is meant to offer a top level message which can be serialized (or encoded, the way to do this in Golgang, if not gRPC, is the gob pacakge offered out of the box as a base package within the lanugage) and passed around, enabling any party involved to understand what this message is about and how to decode it.

@@

\textit{AskForLine} is a message sent from an actor to the city simulator in which the actor asks about the density of any line. This way any actor can take conscious decisions on which route to go on, based on what lies ahead in terms of street densities. When a first route is chosen, between two desired points, the actor can ask about each line which is part of that route. The city simulator will respond with the known density of it. If the actor desires, it can try to find another route by asking GraphHopper service to compute a new route but with an additional rule: avoid a certain point (street, junction, etc.). \textit{RespondWithLine} is the type of message which the city simulator sends back to an actor after it received an \textit{AskForLine} message.

@@

As with every software project started from scratch there are a number of choices to make when choosing software stacks, programming languages, networking models, etc. This section is aiming to explore the rationales behind some of those decisions and how they influenced the building of the current prototype.

@@

Go routines are a lightweight thread of executing. Being lightweight it means they can be easily started, with no overhead and especially without the ceremony of dealing with threads which other popular languages (e.g. Java) have. Listing \ref{lst:goroutine} shows such a routine. It is a part of logic where the city actor responds to either two events: the timer has expired and a decision needs to be made or an information about the line is currently traveling on has been received from the city simulator. Another Go trait is the channels, an indexed communication pipe where there are asynchronous writers and readers, used to decouple two routines. Listing \ref{lst:goroutine} also shows couple of channels. Variables \textit{ticker} and \textit{lineChan} are two channels. They are declared using a type and after that they are used to write and read objects (or structs) of the respective type. The \textit{select} keyword acts like a switch, not on variables and their values but on channels. It will execute the block of code for the channel which has something new to deliver. All those three concepts, \textit{goroutines}, \textit{channels} and \textit{select} switches, combined together makes it easy to write code which takes advantage of multithreading paradigms and which has to run in a heavy networking environment.

@@

The term orchestration means the handling of containers and microservices in order to bring coherence into their interaction and an unifying experience to the end user of a service or of a product. They migth run on multiple nodes (computers or virtual machines) and at the same time they need to communicate with one another. They need to be updated in place and without service disruption. Whenever one of them crash they have to restart as quickly as possible. The services have to be able to discover themselves without dealing with intricancies of IP addresses, proxies, etc. Those are just few concerns around orchestration and the current paper is not aiming to provide a comprehensive view of what it means and what can be achieved with it but rather to set a basis of understanding enough context for the subject of simulating a city with its traffic.

@@

The city simulator is, currently, a single service which means it runs as a single instance as viewed by a city actor. The city actor container runs in multiple instances and it has the need to do so as part of the entire simulation workload. The other two instances which run in the simulation are the routing service, a Graphhopper instance, and the front end web erver which displays a web page for offering visual clues about how the simulation is running. While they run as such there should be an easy way, without friction and additional compute logic to "discover" a certain service. A city actor will need to connect itself to the routing service and to the city simulator. At the same time the front end instance need to connect itself to the city simulator to source its data.

@@

The \textit{docker-compose} tool gives the ability to manipulate more containers and services, with a simple file written in yaml format. Listing \ref{lst:dockercompose} shows the docker-compose file in its brevity and briefness. Let's dissect it. There are four services: \textit{graphhopper}, \textit{citysim}, \textit{cityactor} and \textit{cityfront}. Each of them need an image to run and this image can be specified in two ways: either as an image already compiled and hosted on an container registry or as a local folder which contains a file to build one (usually named \textit{dockerfile}). Because Graphhoper, once compiled with desired maps and settings do not need any more development work, it is uploaded to a personal docker registry and taken from there, whenever needed. The other three services are the places where the most development efforts take place therefore they need to be compiled or recompiled each time the entire traffic simulation application starts.

The containers which are running may need to have different properties. First and most important is the container port which needs to be published. The port from inside the container, where a certain process expect a TCP communication is forwarded to the local host port and thus is accessible from outside the Docker internal network. Another detail of the container is its dependency. For example \textit{cityactor} cannot run without textit{Graphhopper} because it doesn't have any place to obtain a route. Therefore it depends on \textit{Graphhopper}, i.e. it waits for GraphHopper to start first. And on \textit{citysim}, of course. The last bit of detail which can be seen here is the policy of restarting a container. By default the policy is set to "No" which means that the container will not be restarted if there is any failure within it and it stops. However, for simulation purposes, as we need a constant stream of city actors to swarm through the city, the policy is set to "Always" which will restart the container after if closes itself, i.e. it finishes its travel.

@@

Enter the DNS (Domain Name System), the backbone of the entire internet. In its simplest description, avoiding many inherent and nitpicking details, it is a dictionary which keeps track of every registered and easy to memorize name, e.g. \textit{en.wikipedia.com} and the IP addresses where it is located. Any client which would like to access such an name will query first the DNS resolvers and then it will proceed to send a message to the obtained IP address. The most notable interaction of a user with the DNS is the address bar of a browser where the user inserts the name of the site they want to access and while writing the address suggestions are displayed by the browser with the help of recommandation engines \citep{risley2001domain}.

When many resources are needed, to serve a great number of users or to support a great number of city actors, a certain service might be so busy with serving data that the compute and memory resources it needs are greater than the underlying hardware is able to support. Therefore it might be located at few addresses at the same time and any client which wants to access it should access the address where there are available compute and memory resources to serve its queries. Ideal would be to abstract or to decouple this information from the client and make it transparent for it. To do this the client needs to know only the service name and a certain "networking" entity should route its request to the appropiate available service. Such a entity could be a DNS authoritative server which, based on the load, availability and latency of the services will route the request to the appropiate node \citep{swildens2006scalable}.

Docker and Kubernetes are doing a similar job. They offer a DNS service and according to the load of each node where a service run, a request is rooted to a node which will be able to respond to it. This logic is completely decoupled from the clients. Listings \ref{lst:gocityclientdial} and \ref{lst:cityfrontclientdial} shows how both the city front and a city actor calls the city simulator. All what they do is to use the address, the service name, which was specified in the docker-compose file (listing \ref{lst:dockercompose}). Docker or Kubernetes will do the actual job to route the request to the appropiate node of the service. One thing to note is the difference between the two. A city client will communicate via TCP/IP using the Go default serialization package while the city front use a WebSocket communication technology (add reference to the design choices paragraph) but both of them are transparently routed by Docker.

@@

Figure \ref{fig:ScreenshotWebInterface} shows a screenshot of the web interface discussed in paragraph \ref{subsection:citysimulator}, showing a simulation with less than 20 actors. The web interface is a simple: a html/js only web page (vanilla js) which connects to the city simulator as can be seen in listing \ref{lst:cityfrontclientdial}. The city simulator will send a notification to the web page server each time the density changes for a certain line. It can be that a new or first actor entered a certain line (a list of coordinates which represent a street) or it can be that one or the last actor left a line. The web page will draw any line with a density greater than 0 on top of the map.

@@

A laptop used for simulating (MacBook Pro, 16 GB RAM, i5 3.1 GHz) is able to run easily 50 city actors. Above a certain threshold the bottleneck, surprisingly, is not the memory or the cpu but the way docker handles network interfaces. For a larger number of ciy actors a better tool must be used (see paragraph \ref{subsec:kubernetes}). Figure \ref{fig:ScreenshotWebInterface2} shows a simulation with 50 actors running at the same time throughout the city.

@@

As with all technicl projects, there are a number of enhancements which can be further developed around the concept of simulating a city traffic through actors running in microservices, as presented in this paper. This section will briefly propose few of those details in the following paragraphs.

@@

A second idea for enhancing the simulation is to make the actors to have a more granular and diverse logic of traversing the city. The solution proposed into this paper is a basic one: an actor has a route between two random points on the map and then it goes on to travel across that route. It can also be made to look for street densities in advance and act accordingly. If a street is too crowded the actor might choose to avoid it and compute a new route around it. A set of actors can be made to always run between same points (to simulate, for example, buses).

A third idea is to add more interactions between actors. It has been discussed that a first interaction is an indirect one: the number of actors on a certain street, a number called density. But there are many more points of interactions. Let's take junctions. The city simulator can model any junction and then it can permit actors to pass or not through it. Then there are pedestrian crossings where different types of actors can meet each other. Or subways access, elevators for them, bus stations, etc.

A fourth idea is to have contained actors. A bus is an actor. A person is an actor. If a person is waiting on a bus station and then it takes a bus, there are two actors, but they are tied together, or contained and are not independent anymore until the actor-person choose to get out of the actor-bus. This means the densities across sidewalkings and bus stations need to be computed accordingly.

A fifth idea of enhancement is not tied to the way simulation is working but to how it is presented. The web page can display different colors based on the density of a line, offering a visual clue on it. Also it can easily show more information when the cursor is over a point of interest (a street, a junction, etc.). Or it can speed up a past simulation and it can display an animation.
